---
title: "Gapminder"
author: "Liz Bageant"
date: "2/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


```{r echo = FALSE}
library(dslabs) #install.packages("dslabs")
library(tidyverse) 
```

<br>

## Data

We use the Gapminder data. The [Gapminder Project](https://www.gapminder.org/about/) project was pioneered by Hans Rosling, who is famous for describing the prosperity of nations over time through famines, wars and other historic events with this beautiful data visualization in his [2006 TED Talk: The best stats you’ve ever seen](https://www.ted.com/talks/hans_rosling_shows_the_best_stats_you_ve_ever_seen):

Let's explore a subset of the data from 2011. 

```{r}
gap2011 <- gapminder %>% 
  as_tibble() %>% 
  filter(year == 2011)

dim(gap2011)
colnames(gap2011)
head(gap2011)

```


<br>

## Life expectancy

#### Explore the distribution of life expectancy data

Histogram: Not very pretty.
```{r echo = FALSE}
ggplot(data = gap2011, aes(x = life_expectancy)) +
  geom_histogram()
```

Density plot: This is nicer. 
```{r echo = FALSE}
ggplot(data = gap2011, aes(x = life_expectancy)) +
  geom_density() +
  geom_vline(xintercept = mean(data = gap2011, aes(x =life_expectancy)))
```

How does the distribution of life expectancy vary by continent?
```{r echo = FALSE}
ggplot(data = gap2011, aes(x = life_expectancy)) +
  geom_density() +
  geom_vline(xintercept = mean(data = gap2011, aes(x =life_expectancy))) +
  facet_wrap(~ continent)
```

#### What is the relationship between life expectancy and fertility?
Where fertility was higher (e.g. AFrica, Oceania), life expectancy was lower and vice versa.

```{r}
ggplot(data = gap2011) +
  geom_point(mapping = aes(x = fertility, y = life_expectancy, color = continent)) +
  geom_smooth(mapping = aes(x = fertility, y = life_expectancy)) 
  
```
